---
title: "`r params$study_name`"
author: "Robert Limeira"
date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  study_name: Linear Regression Analysis for The spermidine acetyltransferase SpeG regulates transcription of the small RNA rprA
output: 
  conflr::confluence_document:
    parent_id: "393218"
    space_key: "~7120204d1e65b567a34862bddeb1134d81acf6"
    type: "page"
    toc: TRUE
    toc_depth: 2
    code_folding: "none"
    update: TRUE
    use_original_size: TRUE
    interactive: FALSE
---

Analysis performed for:

The spermidine acetyltransferase SpeG regulates transcription of the small RNA RprA

Linda I. Hu, Ekaterina V. Filippova, Joseph Dang, Sergii Pshenychnyi, Jiapeng Ruan, OlgaKiryukhina, Wayne F. Anderson, Misty L. Kuhn, Alan J. Wolfe

bioRxiv 462937; doi: <https://doi.org/10.1101/462937>

# Methods Exerpt From Paper

To determine whether experimental results were statistically significant, a linear regression was performed, comparing all experimental groups with their respective vector controls. All of the regressions used were set up as follows: the calculated rprApromoter activity was the response variable, the overexpressed plasmids or mutant were the explanatory variable, and time was a random effect. OD was not included as an effect on activity as it is already used in the calculation of activity. Time as a random effect was chosen based on the question asked: Accounting for the effects of time on activity does the experimental group in question signifi- cantly affect overall rprApromoter activity? The significance threshold was set at 0.05. The open source program R (version 3.3.2) and packages “lmerTest”, “ggplot2”, and “moments” were used to visualize and analyze the data (76,77,78,79).

**Activity = 1 Miller Unit = 1000 x (Abs420 - (1.75 x Abs550)) / (T x V x Abs600) We are then able to disregard OD**

Hypothesis: Treatment groups (plasmids or mutant) have an effect on rprA Promoter activity compared to their respected vector controls.

Note: Although we check for normality of the data during analysis in order to determine the correct test to use for the question asked for the final test selected normal data is not necessary.

# Analysis Details

```{r setup, include=FALSE, warning=FALSE}

rm(list=ls(all=TRUE))
knitr::opts_chunk$set(echo = TRUE, root.dir = 'Z:/Users/rlimeira/Documents/Work/Loyola/R_Scrap/SPG_Project')
options(scipen=999)
spgdata3A<-read.csv("spgdata3A.csv")
spgdata3B<-read.csv("spgdata3B.csv")
spgdata1A<-read.csv("spgdata1A.csv")
spgdata1B<-read.csv("spgdata1B.csv")

```

## Loading Libraries

```{r libraries, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(lmerTest)
library(ggplot2)
library(moments)
```

## Figure 3A

### Taking a look at the data

```{r, 3Agraph, echo=FALSE, fig.retina=TRUE}
head(spgdata3A, n=10L)
tail(spgdata3A, n=10L)
ggplot(spgdata3A, aes(x = OD, y = Activity, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")

ggplot(spgdata3A, aes(x = Time, y = Activity, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")

ggplot(spgdata3A, aes(x = Time, y = OD, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")
```

### Checking data "Activity" for normality

```{r 3Anormality, echo=TRUE, fig.retina=TRUE}
hist(spgdata3A$Activity)
skewness(spgdata3A$Activity)
kurtosis(spgdata3A$Activity)
```

```{r 3Abasic_LM, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.retina=TRUE}
###Let's look at the data with a basic lm.
#Setting Orthogonal Contrast
#ExpvsCtrl<-c(1,1,-2)# Comparing both experimental contrasts with wild type
#pSpeGvsY135A<-c(1, -1, 0)# Comparing both experimental contrasts among each other
#contrasts(spgdata3A$Cell_Type)<-cbind(ExpvsCtrl, pSpeGvsY135A) #binding the contrasts to the dataset

#Setting non-Orthogonal Contrast
SpgvsCtrl<-c(1,0,0)# Comparing Spg with control
Y135AvsCtrl<-c(0, 1, 0)# Comparing Y135A with control

spgdata3A$Cell_Type <- as.factor(spgdata3A$Cell_Type)

contrasts(spgdata3A$Cell_Type)<-cbind(SpgvsCtrl, Y135AvsCtrl) #binding the contrasts to the dataset

basic.lm<-lm(Activity ~ Cell_Type, data=spgdata3A)
summary(basic.lm)
ggplot(spgdata3A, aes(x = Cell_Type, y = Activity)) +
  geom_point() +
  geom_smooth(method = "lm")
plot(basic.lm, which = 1)
plot(basic.lm, which = 2)
```

### Accounting for time as a random effect

```{r 3Amixed.lmer, echo=FALSE, fig.retina=TRUE}
mixed.lmer<-lmer(Activity ~ Cell_Type + (1|Time), data=spgdata3A)
summary(mixed.lmer)
plot(mixed.lmer)
qqnorm(resid(mixed.lmer))
qqline(resid(mixed.lmer))
```

## Figure 3B

### Take a look at the data

```{r, 3Bgraph, echo=FALSE, fig.retina=TRUE}
head(spgdata3B, n=10L)
tail(spgdata3B, n=10L)
ggplot(spgdata3B, aes(x = OD, y = Activity, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")

ggplot(spgdata3B, aes(x = Time, y = Activity, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")

ggplot(spgdata3B, aes(x = Time, y = OD, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")
```

### Checking data "Activity" for normality

```{r 3Bnormality, echo=TRUE, fig.retina=TRUE}
hist(spgdata3B$Activity)
skewness(spgdata3B$Activity)
kurtosis(spgdata3B$Activity)
```

```{r 3Bbasic_LM, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.retina=TRUE}
###Let's look at the data with a basic lm.
#Setting Orthogonal Contrast
#ExpvsCtrl<-c(1,1,-2)# Comparing both experimental contrasts with wild type
#pSpeGvsY135A<-c(1, -1, 0)# Comparing both experimental contrasts among each other
#contrasts(spgdata3B$Cell_Type)<-cbind(ExpvsCtrl, pSpeGvsY135A) #binding the contrasts to the dataset

#Setting non-Orthogonal Contrast
SpgvsCtrl<-c(1,0,0)# Comparing Spg with control
Y135AvsCtrl<-c(0, 1, 0)# Comparing Y135A with control

spgdata3B$Cell_Type <- as.factor(spgdata3B$Cell_Type)

contrasts(spgdata3B$Cell_Type)<-cbind(SpgvsCtrl, Y135AvsCtrl) #binding the contrasts to the dataset

basic.lm<-lm(Activity ~ Cell_Type, data=spgdata3B)
summary(basic.lm)
ggplot(spgdata3B, aes(x = Cell_Type, y = Activity)) +
  geom_point() +
  geom_smooth(method = "lm")
plot(basic.lm, which = 1)
plot(basic.lm, which = 2)
```

### Accounting for time as a random effect

```{r 3Bmixed.lmer, echo=FALSE, fig.retina=TRUE}
mixed.lmer<-lmer(Activity ~ Cell_Type + (1|Time), data=spgdata3B)
summary(mixed.lmer)
plot(mixed.lmer)
qqnorm(resid(mixed.lmer))
qqline(resid(mixed.lmer))
```

## Figure 1A

### Take a look at the data

```{r, 1Agraph, echo=FALSE, fig.retina=TRUE}
head(spgdata1A, n=10L)
tail(spgdata1A, n=10L)
ggplot(spgdata1A, aes(x = OD, y = Activity, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")

ggplot(spgdata1A, aes(x = Time, y = Activity, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")

ggplot(spgdata1A, aes(x = Time, y = OD, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")
```

### Checking data "Activity" for normality

```{r 1Anormality, echo=TRUE, fig.retina=TRUE}
hist(spgdata1A$Activity)
skewness(spgdata1A$Activity)
kurtosis(spgdata1A$Activity)
```

```{r 1Abasic_LM, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.retina=TRUE}
###Let's look at the data with a basic lm.

PspeGPlasmidvsCtrl<-c(1, 0)# Comparing pspeG plasmid with control

spgdata1A$Cell_Type <- as.factor(spgdata1A$Cell_Type)

contrasts(spgdata1A$Cell_Type)<-PspeGPlasmidvsCtrl #binding the contrasts to the dataset

basic.lm<-lm(Activity ~ Cell_Type, data=spgdata1A)
summary(basic.lm)
ggplot(spgdata1A, aes(x = Cell_Type, y = Activity)) +
  geom_point() +
  geom_smooth(method = "lm")
plot(basic.lm, which = 1)
plot(basic.lm, which = 2)
```

### Accounting for time as a random effect

```{r 1Amixed.lmer, echo=FALSE, fig.retina=TRUE}
mixed.lmer<-lmer(Activity ~ Cell_Type + (1|Time), data=spgdata1A)
summary(mixed.lmer)
plot(mixed.lmer)
qqnorm(resid(mixed.lmer))
qqline(resid(mixed.lmer))
```

## Figure 1B

### Take a look at the data

```{r, 1Bgraph, echo=FALSE, fig.width=4, fig.height=2}
head(spgdata1B, n=10L)
tail(spgdata1B, n=10L)
ggplot(spgdata1B, aes(x = OD, y = Activity, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")

ggplot(spgdata1B, aes(x = Time, y = Activity, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")

ggplot(spgdata1B, aes(x = Time, y = OD, colour = Cell_Type)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "right")
```

### Checking data "Activity" for normality

```{r 1Bnormality, echo=TRUE, fig.retina=TRUE}
hist(spgdata1B$Activity)
skewness(spgdata1B$Activity)
kurtosis(spgdata1B$Activity)
```

```{r 1Bbasic_LM, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.retina=TRUE}
#Let's look at the data with a basic lm.

PspeGvsCtrl<-c(1, 0)# Comparing pspeG plasmid with control

spgdata1B$Cell_Type <- as.factor(spgdata1B$Cell_Type)

contrasts(spgdata1B$Cell_Type)<-PspeGvsCtrl #binding the contrasts to the dataset

basic.lm<-lm(Activity ~ Cell_Type, data=spgdata1B)
summary(basic.lm)
ggplot(spgdata1B, aes(x = Cell_Type, y = Activity)) +
  geom_point() +
  geom_smooth(method = "lm")
plot(basic.lm, which = 1)
plot(basic.lm, which = 2)
```

### Accounting for time as a random effect

```{r 1Bmixed.lmer, echo=FALSE, fig.retina=TRUE}
mixed.lmer<-lmer(Activity ~ Cell_Type + (1|Time), data=spgdata1B)
summary(mixed.lmer)
plot(mixed.lmer)
qqnorm(resid(mixed.lmer))
qqline(resid(mixed.lmer))
```

# Conclusion Exerpts From Paper

When SpeG was overexpressed from a plasmid in the reference strain, PrprA activity was reduced compared to the vector control during late exponential growth and during the transition into early stationary phase (OD \>1.0, Fig 1A, linear regression analysis t = -2.553, p = 0.01472). When speGwas deleted, PrprAactivity increased in the isogenic speG mutant compared to its wild-type parent (Fig 1B, linear regression analysis t = 7.750, p = 8.65E-12). Based on these results, we conclude that SpeG inhibits transcription from PrprA.

...

**Fig 3. The effect of overexpressing SpeG or SpeG(Y135A) in WT cells and overexpressing SpeG in the speE mutant on rprApromoter activity.** A. WT cells carrying the PrprA-lacZfusion (AJW3759) were transformed with either pspeG(pCA24n-speG), pspeG(Y135A)(pCA24n-speG(Y135A)), or the VC (pCA24n) and grown in TB7 supplemented with 50 μM IPTG and chloramphenicol to maintain the plasmid. Cell growth and β-galactosidase activity were assayed. The values represent average promoter activity with standard deviations of five independent cultures. B. The isogenic speEmutant was transformed with either pspeG(pCA24n-speG) or the VC (pCA24n) and cell growth and β-galactosidase activity were assayed as described for 3A. The values represent average promoter activity with standard deviations of five independent cultures. Linear regression comparison results on rprApromoter activity were significant for all experimental groups: WT/pspeGversus WT/VC (t = -3.752, p = 0.000282), WT/pspeG(Y135A) versus WT/VC (t = -2.456, p = 0.015623), and speE/pspeGversus speE/VC (t = -3.470, p = 0.000745).
